\cleardoublepage
\phantomsection
% \chapter*{Bibliographie}
\addcontentsline{toc}{chapter}{Bibliographie}
\small
\footnotesize
\begin{thebibliography}{99}

\bibitem{springbootdoc}
Spring Team,
\emph{Spring Boot Reference Documentation},
\url{https://spring.io/projects/spring-boot}, consultée en 2026.

\bibitem{reactdoc}
Meta Open Source,
\emph{React Documentation},
\url{https://react.dev/}, consultée en 2026.

\bibitem{postgresdoc}
PostgreSQL Global Development Group,
\emph{PostgreSQL Documentation},    
\url{https://www.postgresql.org/docs/}, consultée en 2026.

\bibitem{dockerdoc}
Docker Inc.,
\emph{Docker Documentation},
\url{https://docs.docker.com/}, consultée en 2026.

\bibitem{twelvefactor}
A. Wiggins,
\emph{The Twelve-Factor App},
\url{https://12factor.net/}, 2012.

\bibitem{fielding2000}
R. Fielding,
\emph{Architectural Styles and the Design of Network-based Software Architectures},
thèse de doctorat, University of California, Irvine, 2000.

\bibitem{rfc7231}
R. Fielding, J. Reschke,
\emph{Hypertext Transfer Protocol (HTTP/1.1): Semantics and Content},
IETF RFC 7231, 2014.

\bibitem{rfc7519}
M. Jones, J. Bradley, N. Sakimura,
\emph{JSON Web Token (JWT)},
IETF RFC 7519, 2015.

\bibitem{owasp2021}
OWASP Foundation,
\emph{OWASP Top Ten (2021)},
\url{https://owasp.org/www-project-top-ten/}, consultée en 2026.

\bibitem{iso25010}
ISO/IEC 25010:2011,
\emph{Systems and software engineering -- Systems and software Quality Requirements and Evaluation (SQuaRE) -- System and software quality models},
2011.

\bibitem{ieee1012}
IEEE,
\emph{IEEE 1012 -- Standard for System, Software, and Hardware Verification and Validation},
2016.

\bibitem{sommerville2016}
I. Sommerville,
\emph{Software Engineering},
10e éd., Pearson, 2016.

\bibitem{pressman2014}
R. S. Pressman,
\emph{Software Engineering: A Practitioner's Approach},
8e éd., McGraw-Hill, 2014.

\bibitem{booch2005}
G. Booch, J. Rumbaugh, I. Jacobson,
\emph{The Unified Modeling Language User Guide},
2e éd., Addison-Wesley, 2005.

\bibitem{gamma1994}
E. Gamma, R. Helm, R. Johnson, J. Vlissides,
\emph{Design Patterns: Elements of Reusable Object-Oriented Software},
Addison-Wesley, 1994.

\bibitem{fowler2002}
M. Fowler,
\emph{Patterns of Enterprise Application Architecture},
Addison-Wesley, 2002.

\bibitem{bass2021}
L. Bass, P. Clements, R. Kazman,
\emph{Software Architecture in Practice},
4e éd., Addison-Wesley, 2021.

\bibitem{newman2021}
S. Newman,
\emph{Building Microservices},
2e éd., O'Reilly, 2021.

\bibitem{kleppmann2017}
M. Kleppmann,
\emph{Designing Data-Intensive Applications},
O'Reilly, 2017.

\bibitem{date2003}
C. J. Date,
\emph{An Introduction to Database Systems},
8e éd., Addison-Wesley, 2003.

\bibitem{laraveldoc}
Laravel LLC,
\emph{Laravel Documentation},
\url{https://laravel.com/docs}, consultée en 2026.

\bibitem{sanctumdoc}
Laravel LLC,
\emph{Laravel Sanctum (SPA authentication)},
\url{https://laravel.com/docs/sanctum}, consultée en 2026.

\bibitem{vitedoc}
Vite Team,
\emph{Vite Guide},
\url{https://vitejs.dev/guide/}, consultée en 2026.

\end{thebibliography}
